---
title: "Coursera course Statistical Inference project part 1"
author: "Joris Van den Bossche"
date: "1 september 2016"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)
```
# Overview
This assignment has two parts. Part one covers a simulation of the exponential distribution. From the simulated data the mean and variance will be estimated and tested to the theoretical value through t-testing. Then by plotting the distribution of the mean estimates will be compared to the normal distribution. The results were as expected.

# Simulation of the exponential distribution
In this section the exponential distribution will be simulated using the R function "rexp". The rate or lambda parameter is set to 0.2. 10.000 values are simulated; the first six values are shown below.
```{r}
set.seed(10)
simulationExp <- data.frame(draws = rexp(10000, rate = 0.2))
head(simulationExp)
```
Now to construct a plot using ggPlot2, as follows:
```{r, fig.height=2.5}
library(ggplot2)
ggplot(data = simulationExp) +
  geom_histogram(aes(draws))
```
It looks like the exponential distribution, so that is a good start!  

# Sample Mean versus Theoretical Mean
To check if the simulated mean is identical to the theoretical mean, a simulation of 1000 means is collected. Each mean will be calculated from 40 random draws of the exponential distribution:
```{r,  fig.height=2.5}
set.seed(12)
simulatedMeans = NULL
for (i in 1 : 1000) simulatedMeans = c(simulatedMeans, mean(rexp(40, rate = 0.2)))
simulatedMeans <- data.frame(means = simulatedMeans)
ggplot(data = simulatedMeans) +
  geom_histogram(aes(means))+
  geom_vline(xintercept = mean(simulatedMeans$means), col = "blue")
```
Esimating the original distribution mean by calculating the mean of the simulated means, also allready implemented in the plot above as a blue line:
```{r}
mean(simulatedMeans$means)
```
The theoretical mean of an exponential distribution is the inverse of lambda. In this case of lambda equal 0.2, the inverse is 5. The estimated value is close to this. How close can be tested with a p-value, the null hypothesis being the mean equal to 5:
```{r}
t.test(simulatedMeans$means, mu = 5)
```
The results show a p-value of 0.6825. The null hypothesis is not rejected. We can conclude the mean to be equal to the theoretical value 5.

# Sample Variance versus Theoretical Variance
The variance can be simulated the same way as the mean. Setting the same seed will make us work with the same numbers that were generated when calculating the means.
```{r, fig.height=2.5}
set.seed(12)
simulatedVariance = NULL
for (i in 1 : 1000) simulatedVariance = c(simulatedVariance, sd(rexp(40, rate = 0.2))^2)
simulatedVariance <- data.frame(variances = simulatedVariance)
ggplot(data = simulatedVariance) +
  geom_histogram(aes(variances)) +
  geom_vline(xintercept = mean(simulatedVariance$variances), col = "blue")

```
The original distribution variance can be estimated taking the mean of all variances simulated, also allready implemented in the plot above as a blue line:
```{r}
mean(simulatedVariance$variances)
```
The variance for exponential distributions is lambda to the power of minus two. In this case, that results to 25. We can device another p-test as such with null hypothesis that the variance equals 25:
```{r}
t.test(simulatedVariance$variances, mu = 25)
```
The p-value is .4257, so the null hypothesis is likely and will not be rejected.

# Distribution
Going back to the graph about the mean. The variance in estimating the mean is estimated by:
```{r}
sd(simulatedMeans$means)
```
The theoretical value of this is the population variance divided by the amount of draws, in this case 40 draws, meaning:
```{r}
25/40
```
The two distributions are plotted over the graph using the density instead of a histogram and ggplot's stat_function.
```{r, fig.height=2.5}
ggplot(data = simulatedMeans) +
  geom_density(aes(means)) +
  stat_function(fun = dnorm, 
                colour = "red", 
                args = list(mean = mean(simulatedMeans$means), 
                            sd = sd(simulatedMeans$means)))
```
The central limit theorem sais for sufficiëntly high n, the density function will be normally distributed. For n = 40 this is allready a very good approximation. For fun, let's make a plot with 10000 draws of 4000 samples:
```{r, fig.height=2.5}
set.seed(12)
simulatedMeans = NULL
for (i in 1 : 10000) simulatedMeans = c(simulatedMeans, mean(rexp(4000, rate = 0.2)))
simulatedMeans <- data.frame(means = simulatedMeans)
ggplot(data = simulatedMeans) +
  geom_density(aes(means)) +
  stat_function(fun = dnorm, 
                colour = "red", 
                args = list(mean = mean(simulatedMeans$means), 
                            sd = sd(simulatedMeans$means)))
```
The means is still 5, but the standard error has changed from 6 to about 0.6, divided by 10 as expected when making your sample 100 times bigger.